asc_file_database

### 2019-03-26 src/asc_file_database.Rmd
all raster stack links done
asc outputting

```{r pre-install packages, message=FALSE}
packages <- c("tidyr")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") 
#this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

@@

```

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCLAYER ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2103)

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

>>>>

```{r pre-install packages, message=FALSE}
packages <- c("tidyr", "rJava", "dismo")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r base list or ascii layers}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") 
#this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

@@

write.csv(asclayerall, "../data/env/asc_stacklist.csv", row.names = FALSE)
```

```{r asclist for 1998}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 1999}
#!!!!!!  CHANGE ASCLAYER ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2000}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2001}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2002}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2003}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2004}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2005}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2006}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2007}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2008}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2009}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2010}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2011}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2012}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2013}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

asclayer <- subset(asclayerall, ascyears == 2013)

@@

```{r asclist for 2014}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

@@

```{r asclist for 2015}
#!!!!!!  CHANGE ASCNAME, ASCPATH AND CSV FILE NAME

### 2019-03-26 src/stack_test.Rmd
asc outputting

need tidyr package
```{r}
unyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
unmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
undepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") #this list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

unlayer <- crossing(unyears, unmonths, undepths)
unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used
unlayer$temp_depth <- NA
unlayer$temp_surface <- NA
unlayer$chl_depth <- NA
unlayer$chl_surface <- NA
unlayer$o2_depth <- NA
unlayer$o2_surface <- NA
unlayer$salinity_depth <- NA
unlayer$salinity_surface <- NA
unlayer$nao_sample <- NA
unlayer$nao_prev <- NA
unlayer$nao_winter <- NA
unlayer$amo_sample <- NA
unlayer$amo_prev <- NA
unlayer$amo_winter <- NA

ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

#year <- "X1998"
#month <- "01"
#depth <- "86.9294"
tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
depthvar <- "depth"
surfacevar <- "surface"
naovar <- "nao"
amovar <- "amo"
samplevar <- "sample"
wintervar <- "winter"

for (k in 1:nrow(unlayer)){
    print(unlayer$stackname[k])
    year <- unlayer$unyears[k]
    month <- unlayer$unmonths[k]
    depth <- unlayer$undepths[k]
    for (i in 1:length(ascname)){

    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 5))
    ascdepth <- gsub(".asc", "", ascdepth)
    ascvar1 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
    ascvar2 <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
    ascfile <- ascpath[[i]]
    print(ascfile)

    if (ascyear == year & ascmonth == month & ascvar1 == tempvar & ascvar2 == depthvar & ascdepth == depth){
    unlayer$temp_depth[k] <- ascfile
  } else if (ascyear == year & ascmonth == month & ascvar1 == chlvar & ascvar2 == surfacevar & ascdepth == depth){
   unlayer$chl_surface[k] <- ascfile
 } else if (ascyear == year & ascmonth == month & ascvar1 == o2var & ascvar2 == depthvar & ascdepth == depth){
  unlayer$o2_depth[k] <- ascfile
 } else if (ascyear == year & ascmonth == month & ascvar1 == salvar & ascvar2 == depthvar & ascdepth == depth){
  unlayer$salinity_depth[k] <- ascfile
 } else if (ascyear == year & ascvar1 == naovar & ascvar2 == wintervar){
  unlayer$nao_winter[k] <- ascfile
 } else if (ascyear == year & ascmonth == month  & ascvar1 == naovar & ascvar2 == samplevar){
  unlayer$nao_sample[k] <- ascfile
 }

    }
    write.csv(unlayer, "../data/env/asc_layerlist.csv", row.names = TRUE)
}

@@

@@

  preplot <- plot(pred, png(paste0(stackn, ".png")))

```{r}
var <- raster(stacktest$chl[1])
var
a <- raster(var)
a
```

@@

```{r}
unyears <- as.character(c("1992", "1998")) 
unmonths <- c("01", "03")
undepths <- c("10", "86.9294")
unlayer <- crossing(unyears, unmonths, undepths)
unlayer$stackname <- paste0("st", unlayer$unyears, unlayer$unmonths, unlayer$undepths) #this creates the name of the stack that will be used

unlayer$temp_depth <- NA
unlayer$temp_surface <- NA
unlayer$chl_depth <- NA
unlayer$chl_surface <- NA
unlayer$o2_depth <- NA
unlayer$o2_surface <- NA
unlayer$salinity_depth <- NA
unlayer$salinity_surface <- NA
unlayer$nao_sample <- NA
unlayer$nao_prev <- NA
unlayer$nao_winter <- NA
unlayer$amo_sample <- NA
unlayer$amo_prev <- NA
unlayer$amo_winter <- NA

tempvar <- "temp"
chlvar <- "chl"
o2var <- "o2"
salvar <- "salinity"
depthvar <- "depth"
surfacevar <- "0.50576"
naovar <- "nao"
amovar <- "amo"
samplevar <- "sample"
wintervar <- "winter"

ascname <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = FALSE)    # this folder currently contains other dummy files. its just a test to automate stacks
ascpath <- list.files("../data/jan_test_86.9/", pattern = "*.asc", full.names = TRUE)    # this folder currently contains other dummy files. its just a test to automate stacks

for (k in 1:nrow(unlayer)){
    year <- unlayer$unyears[k]
    month <- unlayer$unmonths[k]
    depth <- unlayer$undepths[k]

    for (i in 1:length(ascname)){
    ascyear <- (sapply(strsplit(ascname[[i]], "_"), "[[", 1))
    ascmonth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 2))
    ascvar <- (sapply(strsplit(ascname[[i]], "_"), "[[", 3))
    ascdepth <- (sapply(strsplit(ascname[[i]], "_"), "[[", 4))
    ascdepth <- gsub(".asc", "", ascdepth)
    ascfile <- ascpath[[i]]
    print(ascfile)
    if (ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == depth){
      unlayer$temp_depth[k] <- ascfile
    } else if(ascyear == year & ascmonth == month & ascvar == tempvar & ascdepth == surfacevar){
        unlayer$temp_surface[i] <- ascfile
    }
    }
}

@@

system.file("java", package="dismo") #finding location to put maxent jar file
system("java -version")
mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))

@@

```{r}
janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
janpres$occ <- 1
janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
janab$occ <- 0
prab <- rbind(janpres, janab)
prab$originalscientificname <- "cap" #because species name needs to be the same
pa <- prab$occ
prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
prab
#write.csv(prab, "../output/bio/maxjan98test.csv", row.names = FALSE)

@@

```{r}
presaball <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
janpresaball <- subset(presaball, month == 1)
janpresaball <- subset(janpresaball, select = c(occurrence, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
paall <- janpresaball$occurrence
janpresaball <- subset(janpresaball, select = -c(occurrence))

@@

with all variables
```{r}
janaball <- read.csv("../output/bio/background_10k.csv", header = TRUE)
janpresall <- read.csv("../output/bio/pres_nomissing.csv", header = TRUE)
janpresaball <- rbind.fill(janaball, janpresall)
janpresaball <- subset(janpresaball, month == 1)
janpresaball <- subset(janpresaball, select = c(occurrence, longitude_meters, latitude_meters, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
names(janpresaball)[names(janpresaball)=="occurrence"] <- "species"
names(janpresaball)[names(janpresaball)=="longitude_meters"] <- "longitude"
names(janpresaball)[names(janpresaball)=="latitude_meters"] <- "latitude"
janpresall <- subset(janpresaball, species == 1)
janpresall$species <- replace(janpresall$species, janpresall$species == "1", "fish")
write.csv(janpresall, "../output/maxentvariableselection/presence.csv", row.names = FALSE)
janaball <- subset(janpresaball, species == 0)
janpresall$species <- replace(janpresall$species, janpresall$species == "0", "bg")
write.csv(janaball, "../output/maxentvariableselection/background.csv", row.names = FALSE)

@@

```{r}
#aic fiddling

>>>>

@@

need tidyr package
```{r}

@@

@@

@@

mod1 <- maxent(x = prab, p = pa, path = paste0("../output/maxentout"), args =  c("responsecurves"), save(md, file = "../output/maxentmod"))

#THIS IS THE GOOD CODE!
unlayer <- read.csv("../data/env/asc_layerlist1998.csv", header = TRUE)
unlayer <- subset(unlayer, ascmonths == 1)
unlayer <- unlayer[-c(1), ]
unlayer[] <- lapply(unlayer, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
aea <- raster("../output/env/aea.tif")

@@

```{r fiddling with maxent args}
janpres <- read.csv("../output/bio/maxenttestjanpr.csv", header = TRUE)
janpres$occ <- 1
janab <- read.csv("../output/bio/maxenttestjanab.csv", header = TRUE)
janab$occ <- 0
prab <- rbind(janpres, janab)
prab$originalscientificname <- "cap" #because species name needs to be the same
pa <- prab$occ
prab <- subset(prab, select = -c(x, y, occ, originalscientificname)) #ok so for maxent in r just need the variables
prab
#write.csv(prab, "../output/bio/maxjan98test.csv", row.names = FALSE)

@@

```{r maxent with all variables}
presaball <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
janpresaball <- subset(presaball, month == 1)
janpresaball <- subset(janpresaball, select = c(occurrence, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
paall <- janpresaball$occurrence
janpresaball <- subset(janpresaball, select = -c(occurrence))

@@

with all variables
```{r maxent variable selection package - not used}
janaball <- read.csv("../output/bio/background_10k.csv", header = TRUE)
janpresall <- read.csv("../output/bio/pres_nomissing.csv", header = TRUE)
janpresaball <- rbind.fill(janaball, janpresall)
janpresaball <- subset(janpresaball, month == 1)
janpresaball <- subset(janpresaball, select = c(occurrence, longitude_meters, latitude_meters, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
names(janpresaball)[names(janpresaball)=="occurrence"] <- "species"
names(janpresaball)[names(janpresaball)=="longitude_meters"] <- "longitude"
names(janpresaball)[names(janpresaball)=="latitude_meters"] <- "latitude"
janpresall <- subset(janpresaball, species == 1)
janpresall$species <- replace(janpresall$species, janpresall$species == "1", "fish")
write.csv(janpresall, "../output/maxentvariableselection/presence.csv", row.names = FALSE)
janaball <- subset(janpresaball, species == 0)
janpresall$species <- replace(janpresall$species, janpresall$species == "0", "bg")
write.csv(janaball, "../output/maxentvariableselection/background.csv", row.names = FALSE)

@@

```{r AIC for regularization selection}
#aic fiddling
